\section{Timeline}

\subsection{Phase 1: Foundation and Dataset Creation (Weeks 1-8)}
\begin{itemize}
    \item \textbf{Weeks 1-2}: Literature review and finalization of anchor-conditioned model architecture.
    \item \textbf{Weeks 3-5}: Development of the anchor-based dataset generation pipeline.
    \item \textbf{Weeks 6-8}: Generation and validation of the anchor-based instructional dataset.
\end{itemize}

\subsection{Phase 2: Model Fine-Tuning and Development (Weeks 9-20)}
\begin{itemize}
    \item \textbf{Weeks 9-12}: Implementation of the data loading and pre-processing for the anchor-target model.
    \item \textbf{Weeks 13-16}: Initial model fine-tuning (LoRA/QLoRA) on the new dataset.
    \item \textbf{Weeks 17-20}: Iterative refinement of the model based on initial results; development of baseline models for ablation.
\end{itemize}

\subsection{Phase 3: Evaluation and Analysis (Weeks 21-36)}
\begin{itemize}
    \item \textbf{Weeks 21-24}: Development of the evaluation protocol, including surveys for the human study and automated metric scripts.
    \item \textbf{Weeks 25-30}: Recruitment of audio professionals and execution of the subjective evaluation study.
    \item \textbf{Weeks 31-36}: Analysis of quantitative and qualitative data; correlation analysis between subjective and automated metrics.
\end{itemize}

\subsection{Phase 4: Dissemination and Finalization (Weeks 37-52)}
\begin{itemize}
    \item \textbf{Weeks 37-44}: Preparation of manuscripts for conference and journal submission based on research findings.
    \item \textbf{Weeks 45-48}: Refinement of open-source code, models, and dataset for public release.
    \item \textbf{Weeks 49-52}: Finalization of thesis/dissertation and dissemination of results through presentations.
\end{itemize}

\subsection{Milestones and Deliverables}
\begin{itemize}
    \item \textbf{Week 8}: Finalized anchor-based dataset.
    \item \textbf{Week 20}: First fine-tuned prototype of the anchor-conditioned model.
    \item \textbf{Week 36}: Completed analysis of human evaluation study and metric correlation.
    \item \textbf{Week 44}: First draft of primary research paper.
    \item \textbf{Week 52}: Final deliverables, including open-source code and pre-trained models.
\end{itemize}

\subsection{Risk Mitigation}
\begin{itemize}
    \item \textbf{Data Quality}: Validate the generated dataset with a small pilot study before full-scale training.
    \item \textbf{Recruitment Challenges}: Begin recruitment for the human evaluation study early and leverage professional networks.
    \item \textbf{Technical Challenges}: Maintain clear versioning of models and experiments; rely on proven PEFT techniques.
    \item \textbf{Timeline Delays}: Incorporate buffer time within each phase and maintain a clear focus on the primary research question.
\end{itemize}
